---
title: "Hw2"
author: 'IRENE MARTINEZ MORATA (UNI: im2557)'
date: "10/5/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
getwd()
```

## Problem 1.1

```{r Problem 1.1}
problem1 <- read_excel("Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
                       sheet = "Mr. Trash Wheel",
                       range = "A2:N534",
                       ) %>%
janitor::clean_names() %>% 
drop_na(dumpster) %>% 
mutate(sports_balls = round(sports_balls, 0))

problem1a <- problem1 %>% 
  filter(year == "2019")

       
```

## Problem 1.2

```{r Problem 1.2}
prec18 <- read_excel("Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
                       sheet = "2018 Precipitation",
                       range = "A2:B14") %>%
  janitor::clean_names() %>% 
  mutate(month = as.numeric(month),
         total = as.numeric(total)) %>% 
   mutate(year = 2018)


prec19 <- read_excel("Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
                       sheet = "2019 Precipitation",
                       range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  mutate(month = as.numeric(month), 
         total = as.numeric(total)) %>%
  drop_na() %>% 
  mutate(year = 2019)

prob1.2 = 
  bind_rows(prec18, prec19) %>% 
  mutate(month = month.name[as.numeric(month)])
```
The Mr. Trash Wheel dataset provides information about the quantity and content of trash in the Mr.Trash Wheel collector, picked up from `r min(pull(problem1, "date"))` to `r max(pull(problem1, "date"))`. 
In the first data set, the wheel collection totals was composed of  composed `r colnames(problem1)`,. The median sports_balls in 2019 was `r median(pull(problem1, "sports_balls"))` 

The second and third data sets contain the precipitation data (in monthly total inches) for the years 2018 and 2019 in Baltimore. There is available data for `r count(prec18)` months, during these the average rainfall in 2018 was `r mean(pull(prec18, "total"))` . For 2019, there is available data for `r count(prec19)` months, and the average rainfall for this year was `r mean(pull(prec19, "total"))`. The cumulative total precipitation during 2018 was `r sum(pull(prec18, "total"))`, and `r sum(pull(prec19, "total"))`, in 2019.


## Problem 2

```{r Problem2, echo=TRUE}
Prob2.1 <- read_csv("pols-month.csv") %>% 
  separate(mon, into = c("year", "month", "day")) %>% 
  mutate(month = month.name[as.numeric(month)]) %>% 
  mutate(president = ifelse(prez_dem == 1, "dem", "gop")) %>% 
  select(-prez_dem, -prez_gop, -day) %>% 
  mutate(year = as.numeric(year)) 


Prob2.2 <- read_csv("snp.csv") %>% 
  separate(date, into = c("month", "day", "year")) %>% 
  mutate(month = month.name[as.numeric(month)]) %>% 
  select(-day) %>% 
  relocate(year, month) %>% 
  mutate(year = case_when(
  year >= 50 ~ paste0("19", year), 
  year <= 15 ~ paste0("20", year)),
  year = as.numeric(year))

Prob2.3 <- read_csv("unemployment.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    jan:dec, 
    names_to = "month",
    values_to = "unemployment_rate"
  ) %>% 
  mutate(
    year = as.numeric(year),
    month = replace(month, month == "jan", "January"),
    month = replace(month, month == "feb", "February"),
    month = replace(month, month == "mar", "March"),
    month = replace(month, month == "apr", "April"),
    month = replace(month, month == "may", "May"),
    month = replace(month, month == "jun", "June"),
    month = replace(month, month == "jul", "July"),
    month = replace(month, month == "aug", "August"),
    month = replace(month, month == "sep", "September"),
    month = replace(month, month == "oct", "October"),
    month = replace(month, month == "nov", "November"),
    month = replace(month, month == "dec", "December"))

  
Full_prob2 <- inner_join(Prob2.1, Prob2.2, by = c("year", "month"))

Full_prob2.1 <- inner_join(Full_prob2, Prob2.3, by = c("year", "month"))


```
The first database contained `r count(Prob2.1)` observations with information about the distribution of representatives, senators and presidents from the republican or democrat parties by month from `r min(pull(Prob2.1, "year"))` until `r max(pull(Prob2.1, "year"))`years. The variables included `r colnames(Prob2.1)`
The second database (coded as Prob2.2) contained `r count(Prob2.2)` observations with information on gross domestic product (GDP) from year `r min(pull(Prob2.2, "year"))`, until `r max(pull(Prob2.2, "year"))` . The mean GDP in such time frame was `r mean(pull(Prob2.2, "close"))`, and the median was `r median(pull(Prob2.2, "close"))`
The third database (coded as Prob2.3) contained `r count(Prob2.3)` observations for unemployment rate at different months from `r min(pull(Prob2.3, "year"))` until `r max(pull(Prob2.3, "year"))` . The mean unemployment rate was `r mean(pull(Prob2.3, "unemployment_rate"))`

In total, after combining the data sets, `r count(Full_prob2.1)` observations had data across all the three data sets, the time frame of the data observations was from `r min(pull(Full_prob2.1, "year"))` until `r max(pull(Full_prob2.1, "year"))` . The set of variables included: `r colnames(Full_prob2.1)`. The median GDP across observations `r median(pull(Full_prob2.1, "close"))` , and the median unemployment rate was `r median(pull(Full_prob2.1, "unemployment_rate"))`

## Problem 3

```{r Problem 3}
Prob3 <- read_csv("Popular_Baby_Names.csv") %>% 
janitor::clean_names() %>% 
  mutate(ethnicity = str_to_lower(ethnicity)) %>% 
  mutate(childs_first_name = str_to_lower(childs_first_name)) %>% 
  mutate(gender = str_to_lower(gender),
         ethnicity = as.character(ethnicity),
         ethnicity = recode(ethnicity, "asian and paci" = "asian and pacific islander", "hispanic" = "hispanic", "black non hisp" = "black non hispanic", "white non hisp" = "white non hispanic")
        )  %>% 
distinct() 

  
Prob3.1 <- Prob3 %>%
  filter(childs_first_name == "olivia" & gender == "female") %>% 
  select(-gender, -childs_first_name, -count) %>% 
pivot_wider(names_from = year_of_birth, values_from = rank)

print(Prob3.1)

Prob3.2 <- Prob3 %>% 
  filter(rank == "1" & gender == "male") %>% 
  select(-gender,-count, -rank) %>% 
  pivot_wider(names_from = year_of_birth, values_from = childs_first_name)

print(Prob3.2)

plot <- Prob3 %>% 
  filter(ethnicity == "white non hispanic" & year_of_birth == "2016") %>% 
  ggplot(aes(x = rank, y = count)) + geom_point()

print(plot)
```
